Dynamique des communautés

### 2025-04-21 Article_BIO500/Article_BIO500.Rmd
Modification Will Article BIO500

---
Titre: Template for preparing your research report submission to PNAS using RMarkdown

# Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
Auteurs:
  - name: "Kyara Boisvert"
    affiliation: "a"
  - name: "William Simard"
    affiliation: "a"
  - name: "Zoé Chol"
    affiliation: "a"

address:
  - code: "a"
    address: "Université de Sherbrooke, Département d'écologie, 2500 Boulevard de l'Université, Sherbrooke, Québec, J1N 3C6"

# For footer text
lead_author_surname: Anonymous

## Remove this if not required    
equal_authors:
  code: 1
  text: "A.O.(Author One) and A.T. (Author Two) contributed equally to this work (remove if not applicable)."

author_contributions: |
  Please provide details of author contributions here.

## Remove this if not required
conflict_of_interest: |
  Please declare any conflict of interest here.

Résumé: |
  Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.

#Introduction
  "À travers le temps, nos écosystèmes se sont formés grâce à une grande diversité d’espèces animales et végétales. Cette richesse spécifique, retrouvé sur Terre, est intrinsèquement et de manière pragmatique essentiel à la survie des êtres humains (Kissling, 2018). Malgré la reconnaissance de ces faits et des efforts de conservations constants au maintient de la biodiversité, elle continue de décliner à un rythme alarmant (Kissling, 2018; Newbold, 2016). Plusieurs populations largement répandues ainsi que des espèces menacées sont en déclins (Tittensor, 2014). Combiné avec l’exploitation humaine des écosystèmes terrestres et des écosystèmes marins, ses facteurs en résultent en un biotope Terrestre qui est utilisé au delà de sa durabilité (Kissling, 2018). À la suite d’analyses statistiques et d’observations d’une banque de données d’une série temporelle comportant des échantillonnages entre les années 1950 et 2020, incluant une variété de taxon végétal et animal, nous en sommes arrivé à trois questions de recherche. La première est : Observe-t-on un déclin de la biodiversité à travers les années? La deuxième question est : Quel taxon a-t-on le plus (ou le moins) observé à travers les années? Puis la dernière est : Quel taxon a le plus décliné depuis les années 1970? Pour répondre à ces questions, nous avons effectuées d’autres analyses statistiques. Elles consistaient l’abondance des différentes espèces retrouvées à travers les années. Avec la répartition de l’abondance des espèces, elle nous permet d’obtenir des preuves sur le niveau de rareté (ou non) d’une espèce particulière selon d’autres espèces (Matthews, 2015)."

Méthode et résultats:
  "Afin de réaliser les analyses statistiques, la banque de données, nommée séries temporelles, a été fourni par Biodiversité Québec. Ces données ont été récoltées sur plusieurs décennies, de 1950 à 2020. Cette banque nous a permis de réaliser plusieurs analyses statistiques. Pour débuter, nous avons importé les données taxonomiques dans le logiciel R. Par la suite, nous avons procédé par un nettoyage de l’information de manière à avoir une base précise. Ensuite, une analyse statistique a été fait dans le but de créer des data frame contenant les éléments, tel que les noms scientifiques, les années, les coordonnées, les sources qui procuraient l’information, etc. Ces data frame ont permis de créer de nouvelles bases de données utilisées dans logiciel SQL de R studio. Puis ces nouveaux data frame ont été utilisés pour créer les figures trois figures suivantes. Le premier graphique a été conçu à l’aide de la fonction “plot”. Le type de lignes, l’épaisseur de la ligne ainsi que la couleur de celle-ci ont été défini selon les caractéristiques désirées. Ce graphique représente la richesse spécifique par année selon le nombre d’espèces observées. La figure deux, est un graphique à bar (fonction “bar plot”) qui représente les différentes classes taxonomiques ainsi que le nombre d’observation sur une échelle logarithmique. Des caractéristiques spécifiques tel que la couleur des bandes et leurs bordures ont été attribués. Et la troisième figure, est aussi un graphique à bar représentant les variations, en pourcentage, des espèces observées selon les différentes classes à partir des années 1970. Les bandes rouges représentent les diminutions en abondances des espèces et les bandes vertes sont les augmentations depuis 1970 selon chaque classes taxonomiques observées."

  "Les analyses ont démontrées, pour la figure 1, qu’il y a eu une augmentation dans la richesse des espèces entre les années 1950 et ~ 1993. De plus, nous observons, sur la figure 1, une forte augmentation du nombre d’espèces observées lors des années 1970. Cependant, nous pouvons aussi voir un déclin dans l’abondance des espèces à partir de l’année 1995. À la suite de cette année, 1995, il y a quelques hausses d’espèces observées, mais de façon générale on remarque un déclin qui s’aggrave d’années en années. Ce qu’on constate pour la figure 2, sont les observations par classes taxonomiques sur une échelle logarithmique. La classe qui a le plus grand nombre total d’observations est les Teleosteis. En revanche, celle qui en a le moins est les Petromyzontis. Pour la troisième figure, on observe les variations, en pourcentage, des différentes classes taxonomiques depuis les années 1970. Les bandes vertes sont des valeurs positives démontrant une augmentation d’espèces de cette classe et les bandes rouges sont des diminutions dans le nombre d’espèces observées, donc des valeurs négatives. Pour cette figure, la classe des Amphibia est celle qui a la plus grande variation (500%). Les classes Chondrostei et Equisetopsida ont une valeur respective de 100%. Dans les valeurs négatives, on constate que la classe Aves est celle qui a la plus grande diminution avec une valeur de - 99%. Nous pouvons aussi observer une valeur de 0 pour la classe des Petromyzontis."
Discussion:

Bibliographie:

@@

This PNAS journal template is provided to help you write your work in the correct journal format. Instructions for use are provided below.

Note: please start your introduction without including the word "Introduction" as a section heading (except for math articles in the Physical Sciences section); this heading is implied in the first paragraphs.

# Guide to using this template {.unnumbered}

Please note that whilst this template provides a preview of the typeset manuscript for submission, to help in this preparation, it will not necessarily be the final publication layout. For more detailed information please see the [PNAS Information for Authors](http://www.pnas.org/site/authors/format.xhtml).

## Author Affiliations {#author-affiliations .unnumbered}

Include department, institution, and complete address, with the ZIP/postal code, for each author. Use lower case letters to match authors with institutions, as shown in the example. Authors with an ORCID ID may supply this information at submission.

## Submitting Manuscripts {#submitting-manuscripts .unnumbered}

All authors must submit their articles at [PNAScentral](http://www.pnascentral.org/cgi-bin/main.plex). If you are using Overleaf to write your article, you can use the "Submit to PNAS" option in the top bar of the editor window.

## Format {#format .unnumbered}

Many authors find it useful to organize their manuscripts with the following order of sections; Title, Author Affiliation, Keywords, Abstract, Significance Statement, Results, Discussion, Materials and methods, Acknowledgments, and References. Other orders and headings are permitted.

## Manuscript Length {#manuscript-length .unnumbered}

PNAS generally uses a two-column format averaging 67 characters, including spaces, per line. The maximum length of a Direct Submission research article is six pages and a PNAS PLUS research article is ten pages including all text, spaces, and the number of characters displaced by figures, tables, and equations. When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces) for Direct Submissions and 72,000 characters (including spaces) for PNAS PLUS.

## References {#references .unnumbered}

References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g. @belkin2002using and [@berard1994embedding; @coifman2005geometric]. All references, including for the SI, should be included in the main manuscript file. References appearing in both sections should not be duplicated. SI references included in tables should be included with the main reference section.

## Data Archival {#data-archival .unnumbered}

PNAS must be able to archive the data essential to a published article. Where such archiving is not possible, deposition of data in public databases, such as GenBank, ArrayExpress, Protein Data Bank, Unidata, and others outlined in the Information for Authors, is acceptable.

## Language-Editing Services {#language-editing-services .unnumbered}

Prior to submission, authors who believe their manuscripts would benefit from professional editing are encouraged to use a language-editing service (see list at www.pnas.org/site/authors/language-editing.xhtml). PNAS does not take responsibility for or endorse these services, and their use has no bearing on acceptance of a manuscript for publication.

![Placeholder image of a frog with a long example caption to show justification setting.](frog.png)

## Digital Figures {#sec:figures .unnumbered}

Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text, and images must be final size. Authors may submit U3D or PRC files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format. Color images must be in RGB (red, green, blue) mode. Include the font files for any text.

Figures and Tables should be labelled and referenced in the standard way using the `\label{}` and `\ref{}` commands.

Figure $$fig:frog$$ shows an example of how to insert a column-wide figure. To insert a figure wider than one column, please use the `\begin{figure*}...\end{figure*}` environment. Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.

## Single column equations {#single-column-equations .unnumbered}

Authors may use 1- or 2-column equations in their article, according to their preference.

To allow an equation to span both columns, options are to use the `\begin{figure*}...\end{figure*}` environment mentioned above for figures, or to use the `\begin{widetext}...\end{widetext}` environment as shown in equation $$eqn:example$$ below.

Please note that this option may run into problems with floats and footnotes, as mentioned in the [cuted package documentation](http://texdoc.net/pkg/cuted). In the case of problems with footnotes, it may be possible to correct the situation using commands `\footnotemark` and `\footnotetext`.

$$\begin{aligned}
(x+y)^3&=(x+y)(x+y)^2\\
       &=(x+y)(x^2+2xy+y^2) \label{eqn:example} \\
       &=x^3+3x^2y+3xy^3+x^3. 
\end{aligned}$$

```{=html}
<!-- pandoc writes all tables using longtable, which fails in 2-column mode

  Species                    CBS     CV     G3
  ----------------------- ------ ------ ------
  1\. Acetaldehyde           0.0    0.0    0.0
  2\. Vinyl alcohol          9.1    9.6   13.5
  3\. Hydroxyethylidene     50.8   51.2   54.0

  : Comparison of the fitted potential energy surfaces and ab initio
  benchmark electronic energy calculations

-->
```
## Supporting Information (SI) {#supporting-information-si .unnumbered}

The main text of the paper must stand on its own without the SI. Refer to SI in the manuscript at an appropriate point in the text. Number supporting figures and tables starting with S1, S2, etc. Authors are limited to no more than 10 SI files, not including movie files. Authors who place detailed materials and methods in SI must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the online methods. If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text. Because PNAS edits SI and composes it into a single PDF, authors must provide the following file formats only.

### SI Text {#si-text .unnumbered}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference).

### SI Figures {#si-figures .unnumbered}

Provide a brief legend for each supporting figure after the supporting text. Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or GIF format; figures may not be embedded in manuscript text. When saving TIFF files, use only LZW compression; do not use JPEG compression. Do not save figure numbers, legends, or author names as part of the image. Composite figures must be pre-assembled.

### 3D Figures {#d-figures .unnumbered}

Supply a composable U3D or PRC file so that it may be edited and composed. Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.

### SI Tables {#si-tables .unnumbered}

Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file. Do not use tabs or spaces to separate columns in Word tables.

### SI Datasets {#si-datasets .unnumbered}

Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.

### SI Movies {#si-movies .unnumbered}

Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file. All movies should be submitted at the desired reproduction size and length. Movies should be no more than 10 MB in size.

### Still images {#still-images .unnumbered}

Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.

### Appendices {#appendices .unnumbered}

PNAS prefers that authors submit individual source files to ensure readability. If this is not possible, supply a single PDF file that contains all of the SI associated with the paper. This file type will be published in raw format and will not be edited or composed.

>>>>

---
title: "Dynamique des communautés"

# Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author

author:
  - name: "Kyara Boisvert"
    affiliation: "Université de Sherbrooke, Département d'écologie, 2500 Boulevard de l'Université, Sherbrooke, Québec, J1N 3C6" 
  - name: "Zoé Chol"
    affiliation: "Université de Sherbrooke, Département d'écologie, 2500 Boulevard de l'Université, Sherbrooke, Québec, J1N 3C6"
  - name: "William Simard"
    affiliation: "Université de Sherbrooke, Département d'écologie, 2500 Boulevard de l'Université, Sherbrooke, Québec, J1N 3C6"

abstract: |
  Please provide an abstract of no more than 250 words in a single paragraph. Abstracts should explain to the general reader the major contributions of the article. References in the abstract must be cited in full within the abstract itself and cited in the text.

@@

À travers le temps, nos écosystèmes se sont formés grâce à une grande diversité d’espèces animales et végétales. Cette richesse spécifique, retrouvé sur Terre, est intrinsèquement et de manière pragmatique essentiel à la survie des êtres humains (Kissling, 2018). Malgré la reconnaissance de ces faits et des efforts de conservations constants au maintient de la biodiversité, elle continue de décliner à un rythme alarmant (Kissling, 2018; Newbold, 2016). Plusieurs populations largement répandues ainsi que des espèces menacées sont en déclins (Tittensor, 2014). Combiné avec l’exploitation humaine des écosystèmes terrestres et des écosystèmes marins, ses facteurs en résultent en un biotope Terrestre qui est utilisé au delà de sa durabilité (Kissling, 2018). À la suite d’analyses statistiques et d’observations d’une banque de données d’une série temporelle comportant des échantillonnages entre les années 1950 et 2020, incluant une variété de taxon végétal et animal, nous en sommes arrivé à trois questions de recherche. La première est : Observe-t-on un déclin de la biodiversité à travers les années? La deuxième question est : Quel taxon a-t-on le plus (ou le moins) observé à travers les années? Puis la dernière est : Quel taxon a le plus décliné depuis les années 1970? Pour répondre à ces questions, nous avons effectuées d’autres analyses statistiques. Elles consistaient l’abondance des différentes espèces retrouvées à travers les années. Avec la répartition de l’abondance des espèces, elle nous permet d’obtenir des preuves sur le niveau de rareté (ou non) d’une espèce particulière selon d’autres espèces (Matthews, 2015)."

### Méthode:

Afin de réaliser les analyses statistiques, la banque de données, nommée séries temporelles, a été fourni par Biodiversité Québec. Ces données ont été récoltées sur plusieurs décennies, de 1950 à 2020. Cette banque nous a permis de réaliser plusieurs analyses statistiques. Pour débuter, nous avons importé les données taxonomiques dans le logiciel R. Par la suite, nous avons procédé par un nettoyage de l’information de manière à avoir une base précise. Ensuite, une analyse statistique a été fait dans le but de créer des data frame contenant les éléments, tel que les noms scientifiques, les années, les coordonnées, les sources qui procuraient l’information, etc. Ces data frame ont permis de créer de nouvelles bases de données utilisées dans logiciel SQL de R studio. Puis ces nouveaux data frame ont été utilisés pour créer les figures trois figures suivantes. Le premier graphique a été conçu à l’aide de la fonction “plot”. Le type de lignes, l’épaisseur de la ligne ainsi que la couleur de celle-ci ont été défini selon les caractéristiques désirées. Ce graphique représente la richesse spécifique par année selon le nombre d’espèces observées. La figure deux, est un graphique à bar (fonction “bar plot”) qui représente les différentes classes taxonomiques ainsi que le nombre d’observation sur une échelle logarithmique. Des caractéristiques spécifiques tel que la couleur des bandes et leurs bordures ont été attribués. Et la troisième figure, est aussi un graphique à bar représentant les variations, en pourcentage, des espèces observées selon les différentes classes à partir des années 1970. Les bandes rouges représentent les diminutions en abondances des espèces et les bandes vertes sont les augmentations depuis 1970 selon chaque classes taxonomiques observées."

### Résultats:

Les analyses ont démontré, pour la figure 1, qu’il y a eu une augmentation dans la richesse des espèces entre les années 1950 et ~ 1993. De plus, nous observons, sur la figure 1, une forte augmentation du nombre d’espèces observées lors des années 1970. Cependant, nous pouvons aussi voir un déclin dans l’abondance des espèces à partir de l’année 1995. À la suite de cette année, 1995, il y a quelques hausses d’espèces observées, mais de façon générale on remarque un déclin qui s’aggrave d’années en années. Ce qu’on constate pour la figure 2, sont les observations par classes taxonomiques sur une échelle logarithmique. La classe qui a le plus grand nombre total d’observations est les Teleosteis. En revanche, celle qui en a le moins est les Petromyzontis. Pour la troisième figure, on observe les variations, en pourcentage, des différentes classes taxonomiques depuis les années 1970. Les bandes vertes sont des valeurs positives démontrant une augmentation d’espèces de cette classe et les bandes rouges sont des diminutions dans le nombre d’espèces observées, donc des valeurs négatives. Pour cette figure, la classe des Amphibia est celle qui a la plus grande variation (500%). Les classes Chondrostei et Equisetopsida ont une valeur respective de 100%. Dans les valeurs négatives, on constate que la classe Aves est celle qui a la plus grande diminution avec une valeur de - 99%. Nous pouvons aussi observer une valeur de 0 pour la classe des Petromyzontis."

### Discussion:

### Conclusion

### Bibliographie:
This PNAS journal template is provided to help you write your work in the correct journal format. Instructions for use are provided below.
